---
title: "ortiz_esm229_ps1"
author: "Elena Ortiz"
date: "1/8/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Part 0: Program file directory, initial conditions, parameters 
#### Part 0a: Set file directory, clear workspace 
```{r}
 #set your own directory here 
    setwd('~/Documents/UCSB/W22/Econ & Policy of CC/PS1/ESM229_PS1')
    rm(list=ls()) # clears workspace
```

   

####Part 0b: Model parameters  
```{r}
# Fixed Economic parameters (LEAVE THIS ALONE FOR NOW)
    p_saving<-c(.22) # output saved, default at 0.22
    p_capShare<-c(.3) # Cobb-Douglas capital share, default at 0.3
    p_capDep<-c(.1) # capital depreciation, default at 0.1
    p_TFP_g<-c(.005) # TFP growth rate, default at 0.005   
    p_damage<-c(.009) # default at 0.009

    # Variable Economic parameters (TRY DIFFERENT VALUES)
    p_emInt_g<-c(-.005) # default at -.005
    p_enInt_g<-c(-.01) # default at -.01
    p_pop_g<-c(0.01) # default at .01
    
    # Fixed climate parameters (LEAVE THIS ALONE)
    p_carDecay<-c(400) # carbon decay, default 400
    p_climDelay<-c(.015) # climate decay, default 0.015

    #Variable climate parameters (TRY DIFFERENT VALUES)
    p_climSens<-c(2.5) # climate sensitivity default 2.5
```

    

###### Part 0c: Initial conditions (in 2010) ###### 

```{r}
# years
    p_years<-c(291)

    # Economic
    ic_pop<-c(6900) # in millions
    ic_enInt<-c(5.98) # in EJ/$trill
    ic_emInt<-c(68.27) # in Mt CO2 /EJ            
    ic_K<-c(139.65) # $trill   
    ic_TFP<-c(.03)

    # Climate
    ic_preCO2<-c(275) # ppm
    ic_nowCO2<-c(380) # ppm
    ic_T<-c(0.8) # degree C above preindustrial

    # creating empty matrices for storing data
    year<-matrix(0,p_years,1)    
    pop<-matrix(0, p_years,1)
    enInt<-matrix(0, p_years,1)
    emInt<-matrix(0, p_years,1)    
    TFP<-matrix(0, p_years,1)
    Y_gross<-matrix(0, p_years,1)
    K<-matrix(0, p_years,1)
    gdppc<-matrix(0, p_years,1)
    CO2ppm<-matrix(0, p_years,1)
    em_MtCO2<-matrix(0,p_years,1)
    em_ppm<-matrix(0,p_years,1)
    Y_net<-matrix(0,p_years,1)
    C<-matrix(0,p_years,1)
    C_pc<-matrix(0,p_years,1)
    Teq<-matrix(0,p_years,1)
    climateDamage<-matrix(0,p_years,1)
```
  
    
###Part 1: economic model 
####Part 1a: Getting exogenous changes in population, energy intensity, emissions intensity, TFP

```{r}
 # initializing for first period    
    year[1]<-2010
    pop[1]<-ic_pop
    enInt[1]<-ic_enInt    
    emInt[1]<-ic_emInt
    TFP[1]<-ic_TFP

    for (j in 2:p_years) {
      year[j]<-year[j-1]+1
      pop[j]=pop[j-1]*(1+p_pop_g)
      enInt[j]=enInt[j-1]*(1+p_enInt_g)
      emInt[j]=emInt[j-1]*(1+p_emInt_g)
      TFP[j]=TFP[j-1]*(1+p_TFP_g)
    }
```

   


###Part 1b: Getting gross output, capital, emissions  

```{r}
 # initializing for first period
    #Part 1: Economic model w/o climate damages
    K[1]<-ic_K
    Y_gross[1]<-TFP[1]*K[1]^p_capShare*pop[1]^(1-p_capShare)
      
    #Part 2: climate model
    #CO2 concentration
    CO2ppm[1]<-ic_nowCO2                      
    T[1]=ic_T

    #Part 3: Economic model with climate damages
    #climate damages
    climateShare<-(p_damage*T[1]^2)/(1+p_damage*T[1]^2) # damage function
    climateDamage[1]<-climateShare*Y_gross[1] #climate damages
    Y_net[1]<-Y_gross[1]-climateDamage[1] # net of damages output          
    C[1]<-Y_net[1]-p_saving*Y_net[1]
    C_pc[1]<-(C[1]/pop[1])*1000      


    # looping over rest of the years
    for (j in 2:p_years) {
      
      #Part 1: Economic model w/o climate damages     
      #Economic output 
      K[j]<-(1-p_capDep)*K[j-1]+p_saving*Y_net[j-1]  #capital accumulation
      Y_gross[j]<-TFP[j]*K[j]^p_capShare*pop[j]^(1-p_capShare) #Cobb-Douglas production function
      gdppc[j]=Y_gross[j]/pop[j]

      #emissions
      em_MtCO2[j]<-pop[j]*gdppc[j]*enInt[j]*emInt[j] #Kaya identity
      em_ppm[j]<-em_MtCO2[j]/7810 # convert to atmospheric CO2 concentrations
      
      #Part 2: climate model     
      #CO2 concentration
      CO2ppm[j]<-CO2ppm[j-1]+0.5*((em_ppm[j])- (CO2ppm[j-1]-(ic_preCO2))/(p_carDecay)) #CO2 concentration accumulation                             
    
      #Radiative forcing
      rad_CO2_Wpm2<-5.35*log(CO2ppm[j]/ic_preCO2)

      # Converting to temperature  
      Teq[j]<-rad_CO2_Wpm2*(p_climSens/(5.3*log(2))) #Equilibrium temperature
      T[j]=T[j-1]+p_climDelay*(Teq[j]-T[j-1]) #Transient temperature

      #Part 3: Economic model with climate damages
      climateShare<-(p_damage*T[j]^2)/(1+p_damage*T[j]^2) # damage function
      climateDamage[j]<-climateShare*Y_gross[j] # climate damages
      Y_net[j]<-Y_gross[j]-climateDamage[j] # output net of damages          
      C[j]<-Y_net[j]-p_saving*Y_net[j] # consumption 
      C_pc[j]<-(C[j]/pop[j])*1000 #per capita consumption     
    }
```

   
  


###Part 4: plots


```{r}
par(mfrow=c(3,2), mar=c(4,4,2,2))
plot(year[2:j], em_MtCO2[2:j], ylab="Emissions (MtCO2")
plot(year[2:j], CO2ppm[2:j], ylab="CO2 concentrations (ppm)")
plot(year[2:j], T[2:j], ylab="Transient temperature (C)")
plot(year[2:j], Teq[2:j], ylab="Equilibrium temperature (C)")
plot(year[2:j], Y_net[2:j], ylab="Net output ($ trillion)")
plot(year[2:j], C_pc[2:j], ylab="Consumption per cap ($ thousand)")
```

### *Part 1*

### Plot 1: BAU scenario 

```{r}
par(mfrow=c(3,2), mar=c(4,4,2,2))
plot(year[2:j], em_MtCO2[2:j], ylab="Emissions (MtCO2")
```
### Plot 2: No change in population 

```{r}
# Fixed Economic parameters (LEAVE THIS ALONE FOR NOW)
    p_saving<-c(.22) # output saved, default at 0.22
    p_capShare<-c(.3) # Cobb-Douglas capital share, default at 0.3
    p_capDep<-c(.1) # capital depreciation, default at 0.1
    p_TFP_g<-c(.005) # TFP growth rate, default at 0.005   
    p_damage<-c(.009) # default at 0.009

    # Variable Economic parameters (TRY DIFFERENT VALUES)
    p_emInt_g<-c(-.005) # default at -.005
    p_enInt_g<-c(-.01) # default at -.01
    p_pop_g<-c(0) # default at .01
    
    # Fixed climate parameters (LEAVE THIS ALONE)
    p_carDecay<-c(400) # carbon decay, default 400
    p_climDelay<-c(.015) # climate decay, default 0.015

    #Variable climate parameters (TRY DIFFERENT VALUES)
    p_climSens<-c(2.5) # climate sensitivity default 2.5
    
 # years
    p_years<-c(291)

    # Economic
    ic_pop<-c(6900) # in millions
    ic_enInt<-c(5.98) # in EJ/$trill
    ic_emInt<-c(68.27) # in Mt CO2 /EJ            
    ic_K<-c(139.65) # $trill   
    ic_TFP<-c(.03)

    # Climate
    ic_preCO2<-c(275) # ppm
    ic_nowCO2<-c(380) # ppm
    ic_T<-c(0.8) # degree C above preindustrial

    # creating empty matrices for storing data
    year<-matrix(0,p_years,1)    
    pop<-matrix(0, p_years,1)
    enInt<-matrix(0, p_years,1)
    emInt<-matrix(0, p_years,1)    
    TFP<-matrix(0, p_years,1)
    Y_gross<-matrix(0, p_years,1)
    K<-matrix(0, p_years,1)
    gdppc<-matrix(0, p_years,1)
    CO2ppm<-matrix(0, p_years,1)
    em_MtCO2<-matrix(0,p_years,1)
    em_ppm<-matrix(0,p_years,1)
    Y_net<-matrix(0,p_years,1)
    C<-matrix(0,p_years,1)
    C_pc<-matrix(0,p_years,1)
    Teq<-matrix(0,p_years,1)
    climateDamage<-matrix(0,p_years,1) 
    
# initializing for first period    
    year[1]<-2010
    pop[1]<-ic_pop
    enInt[1]<-ic_enInt    
    emInt[1]<-ic_emInt
    TFP[1]<-ic_TFP

    for (j in 2:p_years) {
      year[j]<-year[j-1]+1
      pop[j]=pop[j-1]*(1+p_pop_g)
      enInt[j]=enInt[j-1]*(1+p_enInt_g)
      emInt[j]=emInt[j-1]*(1+p_emInt_g)
      TFP[j]=TFP[j-1]*(1+p_TFP_g)
    }

     # initializing for first period
    #Part 1: Economic model w/o climate damages
    K[1]<-ic_K
    Y_gross[1]<-TFP[1]*K[1]^p_capShare*pop[1]^(1-p_capShare)
      
    #Part 2: climate model
    #CO2 concentration
    CO2ppm[1]<-ic_nowCO2                      
    T[1]=ic_T

    #Part 3: Economic model with climate damages
    #climate damages
    climateShare<-(p_damage*T[1]^2)/(1+p_damage*T[1]^2) # damage function
    climateDamage[1]<-climateShare*Y_gross[1] #climate damages
    Y_net[1]<-Y_gross[1]-climateDamage[1] # net of damages output          
    C[1]<-Y_net[1]-p_saving*Y_net[1]
    C_pc[1]<-(C[1]/pop[1])*1000      


    # looping over rest of the years
    for (j in 2:p_years) {
      
      #Part 1: Economic model w/o climate damages     
      #Economic output 
      K[j]<-(1-p_capDep)*K[j-1]+p_saving*Y_net[j-1]  #capital accumulation
      Y_gross[j]<-TFP[j]*K[j]^p_capShare*pop[j]^(1-p_capShare) #Cobb-Douglas production function
      gdppc[j]=Y_gross[j]/pop[j]

      #emissions
      em_MtCO2[j]<-pop[j]*gdppc[j]*enInt[j]*emInt[j] #Kaya identity
      em_ppm[j]<-em_MtCO2[j]/7810 # convert to atmospheric CO2 concentrations
      
    } 
#plot2
par(mfrow=c(3,2), mar=c(4,4,2,2))
plot(year[2:j], em_MtCO2[2:j], ylab="Emissions (MtCO2")  
  
```
#### Plot 3: No change in emissions intensity

```{r}
# Fixed Economic parameters (LEAVE THIS ALONE FOR NOW)
    p_saving<-c(.22) # output saved, default at 0.22
    p_capShare<-c(.3) # Cobb-Douglas capital share, default at 0.3
    p_capDep<-c(.1) # capital depreciation, default at 0.1
    p_TFP_g<-c(.005) # TFP growth rate, default at 0.005   
    p_damage<-c(.009) # default at 0.009

    # Variable Economic parameters (TRY DIFFERENT VALUES)
    p_emInt_g<-c(0) # default at -.005
    p_enInt_g<-c(-.01) # default at -.01
    p_pop_g<-c(0.01) # default at .01
    
    # Fixed climate parameters (LEAVE THIS ALONE)
    p_carDecay<-c(400) # carbon decay, default 400
    p_climDelay<-c(.015) # climate decay, default 0.015

    #Variable climate parameters (TRY DIFFERENT VALUES)
    p_climSens<-c(2.5) # climate sensitivity default 2.5
    
 # years
    p_years<-c(291)

    # Economic
    ic_pop<-c(6900) # in millions
    ic_enInt<-c(5.98) # in EJ/$trill
    ic_emInt<-c(68.27) # in Mt CO2 /EJ            
    ic_K<-c(139.65) # $trill   
    ic_TFP<-c(.03)

    # Climate
    ic_preCO2<-c(275) # ppm
    ic_nowCO2<-c(380) # ppm
    ic_T<-c(0.8) # degree C above preindustrial

    # creating empty matrices for storing data
    year<-matrix(0,p_years,1)    
    pop<-matrix(0, p_years,1)
    enInt<-matrix(0, p_years,1)
    emInt<-matrix(0, p_years,1)    
    TFP<-matrix(0, p_years,1)
    Y_gross<-matrix(0, p_years,1)
    K<-matrix(0, p_years,1)
    gdppc<-matrix(0, p_years,1)
    CO2ppm<-matrix(0, p_years,1)
    em_MtCO2<-matrix(0,p_years,1)
    em_ppm<-matrix(0,p_years,1)
    Y_net<-matrix(0,p_years,1)
    C<-matrix(0,p_years,1)
    C_pc<-matrix(0,p_years,1)
    Teq<-matrix(0,p_years,1)
    climateDamage<-matrix(0,p_years,1) 
    
# initializing for first period    
    year[1]<-2010
    pop[1]<-ic_pop
    enInt[1]<-ic_enInt    
    emInt[1]<-ic_emInt
    TFP[1]<-ic_TFP

    for (j in 2:p_years) {
      year[j]<-year[j-1]+1
      pop[j]=pop[j-1]*(1+p_pop_g)
      enInt[j]=enInt[j-1]*(1+p_enInt_g)
      emInt[j]=emInt[j-1]*(1+p_emInt_g)
      TFP[j]=TFP[j-1]*(1+p_TFP_g)
    }

     # initializing for first period
    #Part 1: Economic model w/o climate damages
    K[1]<-ic_K
    Y_gross[1]<-TFP[1]*K[1]^p_capShare*pop[1]^(1-p_capShare)
      
    #Part 2: climate model
    #CO2 concentration
    CO2ppm[1]<-ic_nowCO2                      
    T[1]=ic_T

    #Part 3: Economic model with climate damages
    #climate damages
    climateShare<-(p_damage*T[1]^2)/(1+p_damage*T[1]^2) # damage function
    climateDamage[1]<-climateShare*Y_gross[1] #climate damages
    Y_net[1]<-Y_gross[1]-climateDamage[1] # net of damages output          
    C[1]<-Y_net[1]-p_saving*Y_net[1]
    C_pc[1]<-(C[1]/pop[1])*1000      


    # looping over rest of the years
    for (j in 2:p_years) {
      
      #Part 1: Economic model w/o climate damages     
      #Economic output 
      K[j]<-(1-p_capDep)*K[j-1]+p_saving*Y_net[j-1]  #capital accumulation
      Y_gross[j]<-TFP[j]*K[j]^p_capShare*pop[j]^(1-p_capShare) #Cobb-Douglas production function
      gdppc[j]=Y_gross[j]/pop[j]

      #emissions
      em_MtCO2[j]<-pop[j]*gdppc[j]*enInt[j]*emInt[j] #Kaya identity
      em_ppm[j]<-em_MtCO2[j]/7810 # convert to atmospheric CO2 concentrations
      
    } 
#plot3
par(mfrow=c(3,2), mar=c(4,4,2,2))
plot(year[2:j], em_MtCO2[2:j], ylab="Emissions (MtCO2")  
```
#### Plot 4: No change in energy intensity

```{r}
# Fixed Economic parameters (LEAVE THIS ALONE FOR NOW)
    p_saving<-c(.22) # output saved, default at 0.22
    p_capShare<-c(.3) # Cobb-Douglas capital share, default at 0.3
    p_capDep<-c(.1) # capital depreciation, default at 0.1
    p_TFP_g<-c(.005) # TFP growth rate, default at 0.005   
    p_damage<-c(.009) # default at 0.009

    # Variable Economic parameters (TRY DIFFERENT VALUES)
    p_emInt_g<-c(-.005) # default at -.005
    p_enInt_g<-c(0) # default at -.01
    p_pop_g<-c(0.01) # default at .01
    
    # Fixed climate parameters (LEAVE THIS ALONE)
    p_carDecay<-c(400) # carbon decay, default 400
    p_climDelay<-c(.015) # climate decay, default 0.015

    #Variable climate parameters (TRY DIFFERENT VALUES)
    p_climSens<-c(2.5) # climate sensitivity default 2.5
    
 # years
    p_years<-c(291)

    # Economic
    ic_pop<-c(6900) # in millions
    ic_enInt<-c(5.98) # in EJ/$trill
    ic_emInt<-c(68.27) # in Mt CO2 /EJ            
    ic_K<-c(139.65) # $trill   
    ic_TFP<-c(.03)

    # Climate
    ic_preCO2<-c(275) # ppm
    ic_nowCO2<-c(380) # ppm
    ic_T<-c(0.8) # degree C above preindustrial

    # creating empty matrices for storing data
    year<-matrix(0,p_years,1)    
    pop<-matrix(0, p_years,1)
    enInt<-matrix(0, p_years,1)
    emInt<-matrix(0, p_years,1)    
    TFP<-matrix(0, p_years,1)
    Y_gross<-matrix(0, p_years,1)
    K<-matrix(0, p_years,1)
    gdppc<-matrix(0, p_years,1)
    CO2ppm<-matrix(0, p_years,1)
    em_MtCO2<-matrix(0,p_years,1)
    em_ppm<-matrix(0,p_years,1)
    Y_net<-matrix(0,p_years,1)
    C<-matrix(0,p_years,1)
    C_pc<-matrix(0,p_years,1)
    Teq<-matrix(0,p_years,1)
    climateDamage<-matrix(0,p_years,1) 
    
# initializing for first period    
    year[1]<-2010
    pop[1]<-ic_pop
    enInt[1]<-ic_enInt    
    emInt[1]<-ic_emInt
    TFP[1]<-ic_TFP

    for (j in 2:p_years) {
      year[j]<-year[j-1]+1
      pop[j]=pop[j-1]*(1+p_pop_g)
      enInt[j]=enInt[j-1]*(1+p_enInt_g)
      emInt[j]=emInt[j-1]*(1+p_emInt_g)
      TFP[j]=TFP[j-1]*(1+p_TFP_g)
    }

     # initializing for first period
    #Part 1: Economic model w/o climate damages
    K[1]<-ic_K
    Y_gross[1]<-TFP[1]*K[1]^p_capShare*pop[1]^(1-p_capShare)
      
    #Part 2: climate model
    #CO2 concentration
    CO2ppm[1]<-ic_nowCO2                      
    T[1]=ic_T

    #Part 3: Economic model with climate damages
    #climate damages
    climateShare<-(p_damage*T[1]^2)/(1+p_damage*T[1]^2) # damage function
    climateDamage[1]<-climateShare*Y_gross[1] #climate damages
    Y_net[1]<-Y_gross[1]-climateDamage[1] # net of damages output          
    C[1]<-Y_net[1]-p_saving*Y_net[1]
    C_pc[1]<-(C[1]/pop[1])*1000      


    # looping over rest of the years
    for (j in 2:p_years) {
      
      #Part 1: Economic model w/o climate damages     
      #Economic output 
      K[j]<-(1-p_capDep)*K[j-1]+p_saving*Y_net[j-1]  #capital accumulation
      Y_gross[j]<-TFP[j]*K[j]^p_capShare*pop[j]^(1-p_capShare) #Cobb-Douglas production function
      gdppc[j]=Y_gross[j]/pop[j]

      #emissions
      em_MtCO2[j]<-pop[j]*gdppc[j]*enInt[j]*emInt[j] #Kaya identity
      em_ppm[j]<-em_MtCO2[j]/7810 # convert to atmospheric CO2 concentrations
      
    } 
#plot4
par(mfrow=c(3,2), mar=c(4,4,2,2))
plot(year[2:j], em_MtCO2[2:j], ylab="Emissions (MtCO2")
```
CO2 emissions decrease under scenarios 2 and 3 relative to the BAU scenario. The emissions in scenario 4 appear to be nearly the same as the BAU case.  

### *Part 2*

#### Part i: emissions intensity not exceeding 2 degrees C
```{r}
# Fixed Economic parameters (LEAVE THIS ALONE FOR NOW)
    p_saving<-c(.22) # output saved, default at 0.22
    p_capShare<-c(.3) # Cobb-Douglas capital share, default at 0.3
    p_capDep<-c(.1) # capital depreciation, default at 0.1
    p_TFP_g<-c(.005) # TFP growth rate, default at 0.005   
    p_damage<-c(.009) # default at 0.009

    # Variable Economic parameters (TRY DIFFERENT VALUES)
    p_emInt_g<-c(.019) # default at -.005
    p_enInt_g<-c(-.01) # default at -.01
    p_pop_g<-c(0.01) # default at .01
    
    # Fixed climate parameters (LEAVE THIS ALONE)
    p_carDecay<-c(400) # carbon decay, default 400
    p_climDelay<-c(.015) # climate decay, default 0.015

    #Variable climate parameters (TRY DIFFERENT VALUES)
    p_climSens<-c(2.5) # climate sensitivity default 2.5
    
# years
    p_years<-c(291)

    # Economic
    ic_pop<-c(6900) # in millions
    ic_enInt<-c(5.98) # in EJ/$trill
    ic_emInt<-c(68.27) # in Mt CO2 /EJ            
    ic_K<-c(139.65) # $trill   
    ic_TFP<-c(.03)

    # Climate
    ic_preCO2<-c(275) # ppm
    ic_nowCO2<-c(380) # ppm
    ic_T<-c(0.8) # degree C above preindustrial

    # creating empty matrices for storing data
    year<-matrix(0,p_years,1)    
    pop<-matrix(0, p_years,1)
    enInt<-matrix(0, p_years,1)
    emInt<-matrix(0, p_years,1)    
    TFP<-matrix(0, p_years,1)
    Y_gross<-matrix(0, p_years,1)
    K<-matrix(0, p_years,1)
    gdppc<-matrix(0, p_years,1)
    CO2ppm<-matrix(0, p_years,1)
    em_MtCO2<-matrix(0,p_years,1)
    em_ppm<-matrix(0,p_years,1)
    Y_net<-matrix(0,p_years,1)
    C<-matrix(0,p_years,1)
    C_pc<-matrix(0,p_years,1)
    Teq<-matrix(0,p_years,1)
    climateDamage<-matrix(0,p_years,1)
    

     # initializing for first period    
    year[1]<-2010
    pop[1]<-ic_pop
    enInt[1]<-ic_enInt    
    emInt[1]<-ic_emInt
    TFP[1]<-ic_TFP

    for (j in 2:p_years) {
      year[j]<-year[j-1]+1
      pop[j]=pop[j-1]*(1+p_pop_g)
      enInt[j]=enInt[j-1]*(1+p_enInt_g)
      emInt[j]=emInt[j-1]*(1+p_emInt_g)
      TFP[j]=TFP[j-1]*(1+p_TFP_g)
    }
    
     # initializing for first period
    #Part 1: Economic model w/o climate damages
    K[1]<-ic_K
    Y_gross[1]<-TFP[1]*K[1]^p_capShare*pop[1]^(1-p_capShare)
      
    #Part 2: climate model
    #CO2 concentration
    CO2ppm[1]<-ic_nowCO2                      
    T[1]=ic_T

    #Part 3: Economic model with climate damages
    #climate damages
    climateShare<-(p_damage*T[1]^2)/(1+p_damage*T[1]^2) # damage function
    climateDamage[1]<-climateShare*Y_gross[1] #climate damages
    Y_net[1]<-Y_gross[1]-climateDamage[1] # net of damages output          
    C[1]<-Y_net[1]-p_saving*Y_net[1]
    C_pc[1]<-(C[1]/pop[1])*1000      


    # looping over rest of the years
    for (j in 2:p_years) {
      
      #Part 1: Economic model w/o climate damages     
      #Economic output 
      K[j]<-(1-p_capDep)*K[j-1]+p_saving*Y_net[j-1]  #capital accumulation
      Y_gross[j]<-TFP[j]*K[j]^p_capShare*pop[j]^(1-p_capShare) #Cobb-Douglas production function
      gdppc[j]=Y_gross[j]/pop[j]

      #emissions
      em_MtCO2[j]<-pop[j]*gdppc[j]*enInt[j]*emInt[j] #Kaya identity
      em_ppm[j]<-em_MtCO2[j]/7810 # convert to atmospheric CO2 concentrations
      
      #Part 2: climate model     
      #CO2 concentration
      CO2ppm[j]<-CO2ppm[j-1]+0.5*((em_ppm[j])- (CO2ppm[j-1]-(ic_preCO2))/(p_carDecay)) #CO2 concentration accumulation                             
    
      #Radiative forcing
      rad_CO2_Wpm2<-5.35*log(CO2ppm[j]/ic_preCO2)

      # Converting to temperature  
      Teq[j]<-rad_CO2_Wpm2*(p_climSens/(5.3*log(2))) #Equilibrium temperature
      T[j]=T[j-1]+p_climDelay*(Teq[j]-T[j-1]) #Transient temperature
    } 
```

Plot
```{r}
plot(year[2:j], Teq[2:j], ylab="Equilibrium temperature (C)")
```
The maximum rate of change of emissions intensity is between 0.018-0.019 to avoid exceeding 2 degrees C.  


#### Part ii: energy intensity not exceeding 2 degrees C
```{r}
# Fixed Economic parameters (LEAVE THIS ALONE FOR NOW)
    p_saving<-c(.22) # output saved, default at 0.22
    p_capShare<-c(.3) # Cobb-Douglas capital share, default at 0.3
    p_capDep<-c(.1) # capital depreciation, default at 0.1
    p_TFP_g<-c(.005) # TFP growth rate, default at 0.005   
    p_damage<-c(.009) # default at 0.009

    # Variable Economic parameters (TRY DIFFERENT VALUES)
    p_emInt_g<-c(-.005) # default at -.005
    p_enInt_g<-c(.014) # default at -.01
    p_pop_g<-c(0.01) # default at .01
    
    # Fixed climate parameters (LEAVE THIS ALONE)
    p_carDecay<-c(400) # carbon decay, default 400
    p_climDelay<-c(.015) # climate decay, default 0.015

    #Variable climate parameters (TRY DIFFERENT VALUES)
    p_climSens<-c(2.5) # climate sensitivity default 2.5
    
# years
    p_years<-c(291)

    # Economic
    ic_pop<-c(6900) # in millions
    ic_enInt<-c(5.98) # in EJ/$trill
    ic_emInt<-c(68.27) # in Mt CO2 /EJ            
    ic_K<-c(139.65) # $trill   
    ic_TFP<-c(.03)

    # Climate
    ic_preCO2<-c(275) # ppm
    ic_nowCO2<-c(380) # ppm
    ic_T<-c(0.8) # degree C above preindustrial

    # creating empty matrices for storing data
    year<-matrix(0,p_years,1)    
    pop<-matrix(0, p_years,1)
    enInt<-matrix(0, p_years,1)
    emInt<-matrix(0, p_years,1)    
    TFP<-matrix(0, p_years,1)
    Y_gross<-matrix(0, p_years,1)
    K<-matrix(0, p_years,1)
    gdppc<-matrix(0, p_years,1)
    CO2ppm<-matrix(0, p_years,1)
    em_MtCO2<-matrix(0,p_years,1)
    em_ppm<-matrix(0,p_years,1)
    Y_net<-matrix(0,p_years,1)
    C<-matrix(0,p_years,1)
    C_pc<-matrix(0,p_years,1)
    Teq<-matrix(0,p_years,1)
    climateDamage<-matrix(0,p_years,1)
    

     # initializing for first period    
    year[1]<-2010
    pop[1]<-ic_pop
    enInt[1]<-ic_enInt    
    emInt[1]<-ic_emInt
    TFP[1]<-ic_TFP

    for (j in 2:p_years) {
      year[j]<-year[j-1]+1
      pop[j]=pop[j-1]*(1+p_pop_g)
      enInt[j]=enInt[j-1]*(1+p_enInt_g)
      emInt[j]=emInt[j-1]*(1+p_emInt_g)
      TFP[j]=TFP[j-1]*(1+p_TFP_g)
    }
    
     # initializing for first period
    #Part 1: Economic model w/o climate damages
    K[1]<-ic_K
    Y_gross[1]<-TFP[1]*K[1]^p_capShare*pop[1]^(1-p_capShare)
      
    #Part 2: climate model
    #CO2 concentration
    CO2ppm[1]<-ic_nowCO2                      
    T[1]=ic_T

    #Part 3: Economic model with climate damages
    #climate damages
    climateShare<-(p_damage*T[1]^2)/(1+p_damage*T[1]^2) # damage function
    climateDamage[1]<-climateShare*Y_gross[1] #climate damages
    Y_net[1]<-Y_gross[1]-climateDamage[1] # net of damages output          
    C[1]<-Y_net[1]-p_saving*Y_net[1]
    C_pc[1]<-(C[1]/pop[1])*1000      


    # looping over rest of the years
    for (j in 2:p_years) {
      
      #Part 1: Economic model w/o climate damages     
      #Economic output 
      K[j]<-(1-p_capDep)*K[j-1]+p_saving*Y_net[j-1]  #capital accumulation
      Y_gross[j]<-TFP[j]*K[j]^p_capShare*pop[j]^(1-p_capShare) #Cobb-Douglas production function
      gdppc[j]=Y_gross[j]/pop[j]

      #emissions
      em_MtCO2[j]<-pop[j]*gdppc[j]*enInt[j]*emInt[j] #Kaya identity
      em_ppm[j]<-em_MtCO2[j]/7810 # convert to atmospheric CO2 concentrations
      
      #Part 2: climate model     
      #CO2 concentration
      CO2ppm[j]<-CO2ppm[j-1]+0.5*((em_ppm[j])- (CO2ppm[j-1]-(ic_preCO2))/(p_carDecay)) #CO2 concentration accumulation                             
    
      #Radiative forcing
      rad_CO2_Wpm2<-5.35*log(CO2ppm[j]/ic_preCO2)

      # Converting to temperature  
      Teq[j]<-rad_CO2_Wpm2*(p_climSens/(5.3*log(2))) #Equilibrium temperature
      T[j]=T[j-1]+p_climDelay*(Teq[j]-T[j-1]) #Transient temperature
    }
```

Plot 
```{r}
plot(year[2:j], Teq[2:j], ylab="Equilibrium temperature (C)")
```
The rate of change of energy intensity cannot exceed about 0.013-0.014 in order to keep equilibrium temperatures below 2 degrees C.  

#### Part iii: emissions intensity for lower and upper ECS bounds

```{r}
# Fixed Economic parameters (LEAVE THIS ALONE FOR NOW)
    p_saving<-c(.22) # output saved, default at 0.22
    p_capShare<-c(.3) # Cobb-Douglas capital share, default at 0.3
    p_capDep<-c(.1) # capital depreciation, default at 0.1
    p_TFP_g<-c(.005) # TFP growth rate, default at 0.005   
    p_damage<-c(.009) # default at 0.009

    # Variable Economic parameters (TRY DIFFERENT VALUES)
    p_emInt_g<-c(.0285) # default at -.005
    p_enInt_g<-c(-.01) # default at -.01
    p_pop_g<-c(0.01) # default at .01
    
    # Fixed climate parameters (LEAVE THIS ALONE)
    p_carDecay<-c(400) # carbon decay, default 400
    p_climDelay<-c(.015) # climate decay, default 0.015

    #Variable climate parameters (TRY DIFFERENT VALUES)
    p_climSens<-c(1.5) # climate sensitivity default 2.5 LOWER
    
# years
    p_years<-c(291)

    # Economic
    ic_pop<-c(6900) # in millions
    ic_enInt<-c(5.98) # in EJ/$trill
    ic_emInt<-c(68.27) # in Mt CO2 /EJ            
    ic_K<-c(139.65) # $trill   
    ic_TFP<-c(.03)

    # Climate
    ic_preCO2<-c(275) # ppm
    ic_nowCO2<-c(380) # ppm
    ic_T<-c(0.8) # degree C above preindustrial

    # creating empty matrices for storing data
    year<-matrix(0,p_years,1)    
    pop<-matrix(0, p_years,1)
    enInt<-matrix(0, p_years,1)
    emInt<-matrix(0, p_years,1)    
    TFP<-matrix(0, p_years,1)
    Y_gross<-matrix(0, p_years,1)
    K<-matrix(0, p_years,1)
    gdppc<-matrix(0, p_years,1)
    CO2ppm<-matrix(0, p_years,1)
    em_MtCO2<-matrix(0,p_years,1)
    em_ppm<-matrix(0,p_years,1)
    Y_net<-matrix(0,p_years,1)
    C<-matrix(0,p_years,1)
    C_pc<-matrix(0,p_years,1)
    Teq<-matrix(0,p_years,1)
    climateDamage<-matrix(0,p_years,1)
    

     # initializing for first period    
    year[1]<-2010
    pop[1]<-ic_pop
    enInt[1]<-ic_enInt    
    emInt[1]<-ic_emInt
    TFP[1]<-ic_TFP

    for (j in 2:p_years) {
      year[j]<-year[j-1]+1
      pop[j]=pop[j-1]*(1+p_pop_g)
      enInt[j]=enInt[j-1]*(1+p_enInt_g)
      emInt[j]=emInt[j-1]*(1+p_emInt_g)
      TFP[j]=TFP[j-1]*(1+p_TFP_g)
    }
    
     # initializing for first period
    #Part 1: Economic model w/o climate damages
    K[1]<-ic_K
    Y_gross[1]<-TFP[1]*K[1]^p_capShare*pop[1]^(1-p_capShare)
      
    #Part 2: climate model
    #CO2 concentration
    CO2ppm[1]<-ic_nowCO2                      
    T[1]=ic_T

    #Part 3: Economic model with climate damages
    #climate damages
    climateShare<-(p_damage*T[1]^2)/(1+p_damage*T[1]^2) # damage function
    climateDamage[1]<-climateShare*Y_gross[1] #climate damages
    Y_net[1]<-Y_gross[1]-climateDamage[1] # net of damages output          
    C[1]<-Y_net[1]-p_saving*Y_net[1]
    C_pc[1]<-(C[1]/pop[1])*1000      


    # looping over rest of the years
    for (j in 2:p_years) {
      
      #Part 1: Economic model w/o climate damages     
      #Economic output 
      K[j]<-(1-p_capDep)*K[j-1]+p_saving*Y_net[j-1]  #capital accumulation
      Y_gross[j]<-TFP[j]*K[j]^p_capShare*pop[j]^(1-p_capShare) #Cobb-Douglas production function
      gdppc[j]=Y_gross[j]/pop[j]

      #emissions
      em_MtCO2[j]<-pop[j]*gdppc[j]*enInt[j]*emInt[j] #Kaya identity
      em_ppm[j]<-em_MtCO2[j]/7810 # convert to atmospheric CO2 concentrations
      
      #Part 2: climate model     
      #CO2 concentration
      CO2ppm[j]<-CO2ppm[j-1]+0.5*((em_ppm[j])- (CO2ppm[j-1]-(ic_preCO2))/(p_carDecay)) #CO2 concentration accumulation                             
    
      #Radiative forcing
      rad_CO2_Wpm2<-5.35*log(CO2ppm[j]/ic_preCO2)

      # Converting to temperature  
      Teq[j]<-rad_CO2_Wpm2*(p_climSens/(5.3*log(2))) #Equilibrium temperature
      T[j]=T[j-1]+p_climDelay*(Teq[j]-T[j-1]) #Transient temperature
    }  
```

Lower bounds plot
```{r}
plot(year[2:j], Teq[2:j], ylab="Equilibrium temperature (C)")
```
The maximum rate of change for emissions intensity is about 0.0285 for the lower bounds. 

```{r}
# Fixed Economic parameters (LEAVE THIS ALONE FOR NOW)
    p_saving<-c(.22) # output saved, default at 0.22
    p_capShare<-c(.3) # Cobb-Douglas capital share, default at 0.3
    p_capDep<-c(.1) # capital depreciation, default at 0.1
    p_TFP_g<-c(.005) # TFP growth rate, default at 0.005   
    p_damage<-c(.009) # default at 0.009

    # Variable Economic parameters (TRY DIFFERENT VALUES)
    p_emInt_g<-c(-2.0) # default at -.005
    p_enInt_g<-c(-.01) # default at -.01
    p_pop_g<-c(0.01) # default at .01
    
    # Fixed climate parameters (LEAVE THIS ALONE)
    p_carDecay<-c(400) # carbon decay, default 400
    p_climDelay<-c(.015) # climate decay, default 0.015

    #Variable climate parameters (TRY DIFFERENT VALUES)
    p_climSens<-c(4.5) # climate sensitivity default 2.5 UPPER
    
# years
    p_years<-c(291)

    # Economic
    ic_pop<-c(6900) # in millions
    ic_enInt<-c(5.98) # in EJ/$trill
    ic_emInt<-c(68.27) # in Mt CO2 /EJ            
    ic_K<-c(139.65) # $trill   
    ic_TFP<-c(.03)

    # Climate
    ic_preCO2<-c(275) # ppm
    ic_nowCO2<-c(380) # ppm
    ic_T<-c(0.8) # degree C above preindustrial

    # creating empty matrices for storing data
    year<-matrix(0,p_years,1)    
    pop<-matrix(0, p_years,1)
    enInt<-matrix(0, p_years,1)
    emInt<-matrix(0, p_years,1)    
    TFP<-matrix(0, p_years,1)
    Y_gross<-matrix(0, p_years,1)
    K<-matrix(0, p_years,1)
    gdppc<-matrix(0, p_years,1)
    CO2ppm<-matrix(0, p_years,1)
    em_MtCO2<-matrix(0,p_years,1)
    em_ppm<-matrix(0,p_years,1)
    Y_net<-matrix(0,p_years,1)
    C<-matrix(0,p_years,1)
    C_pc<-matrix(0,p_years,1)
    Teq<-matrix(0,p_years,1)
    climateDamage<-matrix(0,p_years,1)
    

     # initializing for first period    
    year[1]<-2010
    pop[1]<-ic_pop
    enInt[1]<-ic_enInt    
    emInt[1]<-ic_emInt
    TFP[1]<-ic_TFP

    for (j in 2:p_years) {
      year[j]<-year[j-1]+1
      pop[j]=pop[j-1]*(1+p_pop_g)
      enInt[j]=enInt[j-1]*(1+p_enInt_g)
      emInt[j]=emInt[j-1]*(1+p_emInt_g)
      TFP[j]=TFP[j-1]*(1+p_TFP_g)
    }
    
     # initializing for first period
    #Part 1: Economic model w/o climate damages
    K[1]<-ic_K
    Y_gross[1]<-TFP[1]*K[1]^p_capShare*pop[1]^(1-p_capShare)
      
    #Part 2: climate model
    #CO2 concentration
    CO2ppm[1]<-ic_nowCO2                      
    T[1]=ic_T

    #Part 3: Economic model with climate damages
    #climate damages
    climateShare<-(p_damage*T[1]^2)/(1+p_damage*T[1]^2) # damage function
    climateDamage[1]<-climateShare*Y_gross[1] #climate damages
    Y_net[1]<-Y_gross[1]-climateDamage[1] # net of damages output          
    C[1]<-Y_net[1]-p_saving*Y_net[1]
    C_pc[1]<-(C[1]/pop[1])*1000      


    # looping over rest of the years
    for (j in 2:p_years) {
      
      #Part 1: Economic model w/o climate damages     
      #Economic output 
      K[j]<-(1-p_capDep)*K[j-1]+p_saving*Y_net[j-1]  #capital accumulation
      Y_gross[j]<-TFP[j]*K[j]^p_capShare*pop[j]^(1-p_capShare) #Cobb-Douglas production function
      gdppc[j]=Y_gross[j]/pop[j]

      #emissions
      em_MtCO2[j]<-pop[j]*gdppc[j]*enInt[j]*emInt[j] #Kaya identity
      em_ppm[j]<-em_MtCO2[j]/7810 # convert to atmospheric CO2 concentrations
      
      #Part 2: climate model     
      #CO2 concentration
      CO2ppm[j]<-CO2ppm[j-1]+0.5*((em_ppm[j])- (CO2ppm[j-1]-(ic_preCO2))/(p_carDecay)) #CO2 concentration accumulation                             
    
      #Radiative forcing
      rad_CO2_Wpm2<-5.35*log(CO2ppm[j]/ic_preCO2)

      # Converting to temperature  
      Teq[j]<-rad_CO2_Wpm2*(p_climSens/(5.3*log(2))) #Equilibrium temperature
      T[j]=T[j-1]+p_climDelay*(Teq[j]-T[j-1]) #Transient temperature
    }  
```

Upper bounds plot
```{r}
plot(year[2:j], Teq[2:j], ylab="Equilibrium temperature (C)") 
```

About -2.0 but it gets really screwy here...

#### Part iv: energy intensity for lower and upper ECS bounds 


```{r}
# Fixed Economic parameters (LEAVE THIS ALONE FOR NOW)
    p_saving<-c(.22) # output saved, default at 0.22
    p_capShare<-c(.3) # Cobb-Douglas capital share, default at 0.3
    p_capDep<-c(.1) # capital depreciation, default at 0.1
    p_TFP_g<-c(.005) # TFP growth rate, default at 0.005   
    p_damage<-c(.009) # default at 0.009

    # Variable Economic parameters (TRY DIFFERENT VALUES)
    p_emInt_g<-c(-.005) # default at -.005
    p_enInt_g<-c(.023) # default at -.01
    p_pop_g<-c(0.01) # default at .01
    
    # Fixed climate parameters (LEAVE THIS ALONE)
    p_carDecay<-c(400) # carbon decay, default 400
    p_climDelay<-c(.015) # climate decay, default 0.015

    #Variable climate parameters (TRY DIFFERENT VALUES)
    p_climSens<-c(1.5) # climate sensitivity default 2.5 LOWER
    
# years
    p_years<-c(291)

    # Economic
    ic_pop<-c(6900) # in millions
    ic_enInt<-c(5.98) # in EJ/$trill
    ic_emInt<-c(68.27) # in Mt CO2 /EJ            
    ic_K<-c(139.65) # $trill   
    ic_TFP<-c(.03)

    # Climate
    ic_preCO2<-c(275) # ppm
    ic_nowCO2<-c(380) # ppm
    ic_T<-c(0.8) # degree C above preindustrial

    # creating empty matrices for storing data
    year<-matrix(0,p_years,1)    
    pop<-matrix(0, p_years,1)
    enInt<-matrix(0, p_years,1)
    emInt<-matrix(0, p_years,1)    
    TFP<-matrix(0, p_years,1)
    Y_gross<-matrix(0, p_years,1)
    K<-matrix(0, p_years,1)
    gdppc<-matrix(0, p_years,1)
    CO2ppm<-matrix(0, p_years,1)
    em_MtCO2<-matrix(0,p_years,1)
    em_ppm<-matrix(0,p_years,1)
    Y_net<-matrix(0,p_years,1)
    C<-matrix(0,p_years,1)
    C_pc<-matrix(0,p_years,1)
    Teq<-matrix(0,p_years,1)
    climateDamage<-matrix(0,p_years,1)
    

     # initializing for first period    
    year[1]<-2010
    pop[1]<-ic_pop
    enInt[1]<-ic_enInt    
    emInt[1]<-ic_emInt
    TFP[1]<-ic_TFP

    for (j in 2:p_years) {
      year[j]<-year[j-1]+1
      pop[j]=pop[j-1]*(1+p_pop_g)
      enInt[j]=enInt[j-1]*(1+p_enInt_g)
      emInt[j]=emInt[j-1]*(1+p_emInt_g)
      TFP[j]=TFP[j-1]*(1+p_TFP_g)
    }
    
     # initializing for first period
    #Part 1: Economic model w/o climate damages
    K[1]<-ic_K
    Y_gross[1]<-TFP[1]*K[1]^p_capShare*pop[1]^(1-p_capShare)
      
    #Part 2: climate model
    #CO2 concentration
    CO2ppm[1]<-ic_nowCO2                      
    T[1]=ic_T

    #Part 3: Economic model with climate damages
    #climate damages
    climateShare<-(p_damage*T[1]^2)/(1+p_damage*T[1]^2) # damage function
    climateDamage[1]<-climateShare*Y_gross[1] #climate damages
    Y_net[1]<-Y_gross[1]-climateDamage[1] # net of damages output          
    C[1]<-Y_net[1]-p_saving*Y_net[1]
    C_pc[1]<-(C[1]/pop[1])*1000      


    # looping over rest of the years
    for (j in 2:p_years) {
      
      #Part 1: Economic model w/o climate damages     
      #Economic output 
      K[j]<-(1-p_capDep)*K[j-1]+p_saving*Y_net[j-1]  #capital accumulation
      Y_gross[j]<-TFP[j]*K[j]^p_capShare*pop[j]^(1-p_capShare) #Cobb-Douglas production function
      gdppc[j]=Y_gross[j]/pop[j]

      #emissions
      em_MtCO2[j]<-pop[j]*gdppc[j]*enInt[j]*emInt[j] #Kaya identity
      em_ppm[j]<-em_MtCO2[j]/7810 # convert to atmospheric CO2 concentrations
      
      #Part 2: climate model     
      #CO2 concentration
      CO2ppm[j]<-CO2ppm[j-1]+0.5*((em_ppm[j])- (CO2ppm[j-1]-(ic_preCO2))/(p_carDecay)) #CO2 concentration accumulation                             
    
      #Radiative forcing
      rad_CO2_Wpm2<-5.35*log(CO2ppm[j]/ic_preCO2)

      # Converting to temperature  
      Teq[j]<-rad_CO2_Wpm2*(p_climSens/(5.3*log(2))) #Equilibrium temperature
      T[j]=T[j-1]+p_climDelay*(Teq[j]-T[j-1]) #Transient temperature
    }  
```

lower bounds plot
```{r}
plot(year[2:j], Teq[2:j], ylab="Equilibrium temperature (C)")
```
About .023 max rate of change for energy intensity in the lower bounds scenario.

```{r}
# Fixed Economic parameters (LEAVE THIS ALONE FOR NOW)
    p_saving<-c(.22) # output saved, default at 0.22
    p_capShare<-c(.3) # Cobb-Douglas capital share, default at 0.3
    p_capDep<-c(.1) # capital depreciation, default at 0.1
    p_TFP_g<-c(.005) # TFP growth rate, default at 0.005   
    p_damage<-c(.009) # default at 0.009

    # Variable Economic parameters (TRY DIFFERENT VALUES)
    p_emInt_g<-c(-.005) # default at -.005
    p_enInt_g<-c(-1.9) # default at -.01
    p_pop_g<-c(0.01) # default at .01
    
    # Fixed climate parameters (LEAVE THIS ALONE)
    p_carDecay<-c(400) # carbon decay, default 400
    p_climDelay<-c(.015) # climate decay, default 0.015

    #Variable climate parameters (TRY DIFFERENT VALUES)
    p_climSens<-c(4.5) # climate sensitivity default 2.5 UPPER
    
# years
    p_years<-c(291)

    # Economic
    ic_pop<-c(6900) # in millions
    ic_enInt<-c(5.98) # in EJ/$trill
    ic_emInt<-c(68.27) # in Mt CO2 /EJ            
    ic_K<-c(139.65) # $trill   
    ic_TFP<-c(.03)

    # Climate
    ic_preCO2<-c(275) # ppm
    ic_nowCO2<-c(380) # ppm
    ic_T<-c(0.8) # degree C above preindustrial

    # creating empty matrices for storing data
    year<-matrix(0,p_years,1)    
    pop<-matrix(0, p_years,1)
    enInt<-matrix(0, p_years,1)
    emInt<-matrix(0, p_years,1)    
    TFP<-matrix(0, p_years,1)
    Y_gross<-matrix(0, p_years,1)
    K<-matrix(0, p_years,1)
    gdppc<-matrix(0, p_years,1)
    CO2ppm<-matrix(0, p_years,1)
    em_MtCO2<-matrix(0,p_years,1)
    em_ppm<-matrix(0,p_years,1)
    Y_net<-matrix(0,p_years,1)
    C<-matrix(0,p_years,1)
    C_pc<-matrix(0,p_years,1)
    Teq<-matrix(0,p_years,1)
    climateDamage<-matrix(0,p_years,1)
    

     # initializing for first period    
    year[1]<-2010
    pop[1]<-ic_pop
    enInt[1]<-ic_enInt    
    emInt[1]<-ic_emInt
    TFP[1]<-ic_TFP

    for (j in 2:p_years) {
      year[j]<-year[j-1]+1
      pop[j]=pop[j-1]*(1+p_pop_g)
      enInt[j]=enInt[j-1]*(1+p_enInt_g)
      emInt[j]=emInt[j-1]*(1+p_emInt_g)
      TFP[j]=TFP[j-1]*(1+p_TFP_g)
    }
    
     # initializing for first period
    #Part 1: Economic model w/o climate damages
    K[1]<-ic_K
    Y_gross[1]<-TFP[1]*K[1]^p_capShare*pop[1]^(1-p_capShare)
      
    #Part 2: climate model
    #CO2 concentration
    CO2ppm[1]<-ic_nowCO2                      
    T[1]=ic_T

    #Part 3: Economic model with climate damages
    #climate damages
    climateShare<-(p_damage*T[1]^2)/(1+p_damage*T[1]^2) # damage function
    climateDamage[1]<-climateShare*Y_gross[1] #climate damages
    Y_net[1]<-Y_gross[1]-climateDamage[1] # net of damages output          
    C[1]<-Y_net[1]-p_saving*Y_net[1]
    C_pc[1]<-(C[1]/pop[1])*1000      


    # looping over rest of the years
    for (j in 2:p_years) {
      
      #Part 1: Economic model w/o climate damages     
      #Economic output 
      K[j]<-(1-p_capDep)*K[j-1]+p_saving*Y_net[j-1]  #capital accumulation
      Y_gross[j]<-TFP[j]*K[j]^p_capShare*pop[j]^(1-p_capShare) #Cobb-Douglas production function
      gdppc[j]=Y_gross[j]/pop[j]

      #emissions
      em_MtCO2[j]<-pop[j]*gdppc[j]*enInt[j]*emInt[j] #Kaya identity
      em_ppm[j]<-em_MtCO2[j]/7810 # convert to atmospheric CO2 concentrations
      
      #Part 2: climate model     
      #CO2 concentration
      CO2ppm[j]<-CO2ppm[j-1]+0.5*((em_ppm[j])- (CO2ppm[j-1]-(ic_preCO2))/(p_carDecay)) #CO2 concentration accumulation                             
    
      #Radiative forcing
      rad_CO2_Wpm2<-5.35*log(CO2ppm[j]/ic_preCO2)

      # Converting to temperature  
      Teq[j]<-rad_CO2_Wpm2*(p_climSens/(5.3*log(2))) #Equilibrium temperature
      T[j]=T[j-1]+p_climDelay*(Teq[j]-T[j-1]) #Transient temperature
    } 
```

upper bounds plot 

```{r}
plot(year[2:j], Teq[2:j], ylab="Equilibrium temperature (C)")
```
Approximately max. rate of change of -1.9 but it gets wonky, not realistic 

### *Part 3*: Social cost of carbon 

BAU damages 
```{r}
# Fixed Economic parameters (LEAVE THIS ALONE FOR NOW)
    p_saving<-c(.22) # output saved, default at 0.22
    p_capShare<-c(.3) # Cobb-Douglas capital share, default at 0.3
    p_capDep<-c(.1) # capital depreciation, default at 0.1
    p_TFP_g<-c(.005) # TFP growth rate, default at 0.005   
    p_damage<-c(.009) # default at 0.009

    # Variable Economic parameters (TRY DIFFERENT VALUES)
    p_emInt_g<-c(-.005) # default at -.005
    p_enInt_g<-c(-.01) # default at -.01
    p_pop_g<-c(0.01) # default at .01
    
    # Fixed climate parameters (LEAVE THIS ALONE)
    p_carDecay<-c(400) # carbon decay, default 400
    p_climDelay<-c(.015) # climate decay, default 0.015

    #Variable climate parameters (TRY DIFFERENT VALUES)
    p_climSens<-c(2.5) # climate sensitivity default 2.5
```

    

```{r}
# years
    p_years<-c(291)

#initial conditions
    # Economic
    ic_pop<-c(6900) # in millions
    ic_enInt<-c(5.98) # in EJ/$trill
    ic_emInt<-c(68.27) # in Mt CO2 /EJ            
    ic_K<-c(139.65) # $trill   
    ic_TFP<-c(.03)

    # Climate
    ic_preCO2<-c(275) # ppm
    ic_nowCO2<-c(380) # ppm
    ic_T<-c(0.8) # degree C above preindustrial

    # creating empty matrices for storing data
    year<-matrix(0,p_years,1)    
    pop<-matrix(0, p_years,1)
    enInt<-matrix(0, p_years,1)
    emInt<-matrix(0, p_years,1)    
    TFP<-matrix(0, p_years,1)
    Y_gross<-matrix(0, p_years,1)
    K<-matrix(0, p_years,1)
    gdppc<-matrix(0, p_years,1)
    CO2ppm<-matrix(0, p_years,1)
    em_MtCO2<-matrix(0,p_years,1)
    em_ppm<-matrix(0,p_years,1)
    Y_net<-matrix(0,p_years,1)
    C<-matrix(0,p_years,1)
    C_pc<-matrix(0,p_years,1)
    Teq<-matrix(0,p_years,1)
    climateDamage<-matrix(0,p_years,1)
    
    # initializing for first period    
    year[1]<-2010
    pop[1]<-ic_pop
    enInt[1]<-ic_enInt    
    emInt[1]<-ic_emInt
    TFP[1]<-ic_TFP

    for (j in 2:p_years) {
      year[j]<-year[j-1]+1
      pop[j]=pop[j-1]*(1+p_pop_g)
      enInt[j]=enInt[j-1]*(1+p_enInt_g)
      emInt[j]=emInt[j-1]*(1+p_emInt_g)
      TFP[j]=TFP[j-1]*(1+p_TFP_g)
    }
      
       # initializing for first period (BAU)
    #Part 1: Economic model w/o climate damages
    K[1]<-ic_K
    Y_gross[1]<-TFP[1]*K[1]^p_capShare*pop[1]^(1-p_capShare)
      
    #Part 2: climate model
    #CO2 concentration
    CO2ppm[1]<-ic_nowCO2                      
    T[1]=ic_T

    #Part 3: Economic model with climate damages
    #climate damages
    climateShare<-(p_damage*T[1]^2)/(1+p_damage*T[1]^2) # damage function
    climateDamage[1]<-climateShare*Y_gross[1] #climate damages
    Y_net[1]<-Y_gross[1]-climateDamage[1] # net of damages output          
    C[1]<-Y_net[1]-p_saving*Y_net[1]
    C_pc[1]<-(C[1]/pop[1])*1000      


    # looping over rest of the years
    for (j in 2:p_years) {
      
      #Part 1: Economic model w/o climate damages     
      #Economic output 
      K[j]<-(1-p_capDep)*K[j-1]+p_saving*Y_net[j-1]  #capital accumulation
      Y_gross[j]<-TFP[j]*K[j]^p_capShare*pop[j]^(1-p_capShare) #Cobb-Douglas production function
      gdppc[j]=Y_gross[j]/pop[j]

      #emissions
      em_MtCO2[j]<-pop[j]*gdppc[j]*enInt[j]*emInt[j] #Kaya identity
      em_ppm[j]<-em_MtCO2[j]/7810 # convert to atmospheric CO2 concentrations
      
      #Part 2: climate model     
      #CO2 concentration
      CO2ppm[j]<-CO2ppm[j-1]+0.5*((em_ppm[j])- (CO2ppm[j-1]-(ic_preCO2))/(p_carDecay)) #CO2 concentration accumulation                             
    
      #Radiative forcing
      rad_CO2_Wpm2<-5.35*log(CO2ppm[j]/ic_preCO2)

      # Converting to temperature  
      Teq[j]<-rad_CO2_Wpm2*(p_climSens/(5.3*log(2))) #Equilibrium temperature
      T[j]=T[j-1]+p_climDelay*(Teq[j]-T[j-1]) #Transient temperature

      #Part 3: Economic model with climate damages
      climateShare<-(p_damage*T[j]^2)/(1+p_damage*T[j]^2) # damage function
      climateDamage[j]<-climateShare*Y_gross[j] # climate damages
      Y_net[j]<-Y_gross[j]-climateDamage[j] # output net of damages          
      C[j]<-Y_net[j]-p_saving*Y_net[j] # consumption 
      C_pc[j]<-(C[j]/pop[j])*1000 #per capita consumption     
    }
    
   bau_damages <- climateDamage[j]  #save bau annual damages as a vector 
```

Extra emissions scenario 

```{r}
# years
    p_years<-c(291)

    # Economic
    ic_pop<-c(6900) # in millions
    ic_enInt<-c(5.98) # in EJ/$trill
    ic_emInt<-c(68.27) # in Mt CO2 /EJ ADD .000001 here? = 1 ton           
    ic_K<-c(139.65) # $trill   
    ic_TFP<-c(.03)

    # Climate
    ic_preCO2<-c(275) # ppm
    ic_nowCO2<-c(380) # ppm
    ic_T<-c(0.8) # degree C above preindustrial

    # creating empty matrices for storing data
    year<-matrix(0,p_years,1)    
    pop<-matrix(0, p_years,1)
    enInt<-matrix(0, p_years,1)
    emInt<-matrix(0, p_years,1)    
    TFP<-matrix(0, p_years,1)
    Y_gross<-matrix(0, p_years,1)
    K<-matrix(0, p_years,1)
    gdppc<-matrix(0, p_years,1)
    CO2ppm<-matrix(0, p_years,1)
    em_MtCO2<-matrix(0,p_years,1)
    em_ppm<-matrix(0,p_years,1)
    Y_net<-matrix(0,p_years,1)
    C<-matrix(0,p_years,1)
    C_pc<-matrix(0,p_years,1)
    Teq<-matrix(0,p_years,1)
    climateDamage<-matrix(0,p_years,1) 
    extra_em<-matrix(0,p_years,1)
    extra_em[2]<-(0.000001)

# initializing for first period    
    year[1]<-2010
    pop[1]<-ic_pop
    enInt[1]<-ic_enInt    
    emInt[1]<-ic_emInt
    TFP[1]<-ic_TFP
    
# for year 2
    year[2]<-2011
    pop[2]<-ic_pop
    enInt[2]<-ic_enInt    
    emInt[2]<-ic_emInt
    TFP[2]<-ic_TFP
    
    for (j in 3:p_years) {
      year[j]<-year[j-1]+1
      pop[j]=pop[j-1]*(1+p_pop_g)
      enInt[j]=enInt[j-1]*(1+p_enInt_g)
      emInt[j]=emInt[j-1]*(1+p_emInt_g)
      TFP[j]=TFP[j-1]*(1+p_TFP_g)
    }
      
       # initializing for first period (BAU)
    #Part 1: Economic model w/o climate damages
    K[1]<-ic_K
    Y_gross[1]<-TFP[1]*K[1]^p_capShare*pop[1]^(1-p_capShare)
      
    #Part 2: climate model
    #CO2 concentration
    CO2ppm[1]<-ic_nowCO2                      
    T[1]=ic_T

    #Part 3: Economic model with climate damages
    #climate damages
    climateShare<-(p_damage*T[1]^2)/(1+p_damage*T[1]^2) # damage function
    climateDamage[1]<-climateShare*Y_gross[1] #climate damages
    Y_net[1]<-Y_gross[1]-climateDamage[1] # net of damages output          
    C[1]<-Y_net[1]-p_saving*Y_net[1]
    C_pc[1]<-(C[1]/pop[1])*1000      


    # looping over rest of the years
    for (j in 2:p_years) {
      
      #Part 1: Economic model w/o climate damages     
      #Economic output 
      K[j]<-(1-p_capDep)*K[j-1]+p_saving*Y_net[j-1]  #capital accumulation
      Y_gross[j]<-TFP[j]*K[j]^p_capShare*pop[j]^(1-p_capShare) #Cobb-Douglas production function
      gdppc[j]=Y_gross[j]/pop[j]

      #emissions
      em_MtCO2[j]<-pop[j]*gdppc[j]*enInt[j]*emInt[j] + extra_em[j] #Kaya identity: ADDED .000001 here
      em_ppm[j]<-em_MtCO2[j]/7810 # convert to atmospheric CO2 concentrations
      
      #Part 2: climate model     
      #CO2 concentration
      CO2ppm[j]<-CO2ppm[j-1]+0.5*((em_ppm[j])- (CO2ppm[j-1]-(ic_preCO2))/(p_carDecay)) #CO2 concentration accumulation                             
    
      #Radiative forcing
      rad_CO2_Wpm2<-5.35*log(CO2ppm[j]/ic_preCO2)

      # Converting to temperature  
      Teq[j]<-rad_CO2_Wpm2*(p_climSens/(5.3*log(2))) #Equilibrium temperature
      T[j]=T[j-1]+p_climDelay*(Teq[j]-T[j-1]) #Transient temperature

      #Part 3: Economic model with climate damages
      climateShare<-(p_damage*T[j]^2)/(1+p_damage*T[j]^2) # damage function
      climateDamage[j]<-climateShare*Y_gross[j] # climate damages
      Y_net[j]<-Y_gross[j]-climateDamage[j] # output net of damages          
      C[j]<-Y_net[j]-p_saving*Y_net[j] # consumption 
      C_pc[j]<-(C[j]/pop[j])*1000 #per capita consumption     
    }
    
   extra_damages <- Y_net[j]  #save extra damages as a vector 
```

